---
title: "Reshaping & Matplotlib"
subtitle: ""
author: "Andrew Bray"
output: html_document
---

```{r}
library(tidyverse)
```

#### Question 1

Take a look at the following 4 representations of the same data set concerning occurrences of Tuberculosis in various countries.

```{python}
A = r.table3
B = r.table2
C = r.table1
D1 = r.table4a
D2 = r.table4b
```

For each of the representations, what are the *observations* or *cases*? These refer to specific, unique, and similar sorts of things.  What are the *variables*? These should take the same sort of *value* for each case.


#### Question 2

Extend the approach to tidying `D1` that we demonstrated to form `C` from `D1` and `D2`.

```{python}
```

#### Question 3

```{python}
```


Modify the title, labels, and axis limits with

- `ax.set_xlabel()`
- `ax.set_ylabel()`
- `ax.set_xlim()`
- `ax.set_ylim()`
- `ax.set_title()`

#### Question 4

```{r}
library(palmerpenguins)
data(penguins)
```

```{python}
penguins = r.penguins
```

Create a scatter plot of bill length on bill depth by passing each as a series argument to `.plot()`.


