---
title: "Pandas II"
subtitle: "group by"
author: "Andrew Bray"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Agenda

1.  Review
2.  Stuff
3.  Group by

# US Elections Data

This is county-level data from the presidential elections between 2000 and 2016, provided by the MIT Elections Lab.

```{r}
# remotes::install_github("andrewpbray/boxofdata")
library(boxofdata)
library(tidyverse)
data(uselections)
dim(uselections)
names(uselections)
str(uselections)
```

### Bringing the data into Python

We bring the data into Python and do the our Python equivalent of greeing our data.

```{python}
import pandas as pd
uselections = r.uselections
uselections.shape
uselections.columns
uselections.dtypes
```

### Select columns

**Method 1:** Pass a *string* into `[]`...

```{python}
uselections["county"]
```

... get out a series.

**Method 2:** Pass a *list* into `[]`...

```{python}
uselections[["county"]]
```

... get out a data frame.

### Slicing rows

Pass a *slice* into `[]`...

```{python}
uselections[0:5]
```

... get out a slice of a data frame. Sound familiar?

```{r}
slice(uselections, 1:2)
```

### Selecting and slicing

**Method 1:** (preferred) access labels with `.loc`.

```{python}
uselections.loc[0:5, ["county"]]
```

**Method 2:** access integer indices with `.iloc`.

```{python}
uselections.columns
uselections.iloc[0:5, 3]
```

### Filtering rows

You can apply a Boolean series as a mask.

```{python boolrows, eval = FALSE}
mask = uselections["year"] == 2016
uselections[mask]
```

```{python ref.label = "boolrows", echo = FALSE}
```

## Filtering rows and selecting columns

Boolean mask plus a list of columns.

```{python boolandcolsa, eval = FALSE}
mask = uselections["year"].isin([2012, 2016])
uselections[mask, ["county", "state"]]
```

Will this run?

> Need to use `.loc`

```{python boolandcolsb, eval = FALSE}
mask = uselections["year"].isin([2012, 2016])
uselections.loc[mask, ["county", "state"]] #<<
```

```{python ref.label = "boolandcolsb", echo = FALSE}
```

### Let's shine that up.

1.  Form data frame.
2.  Apply `.agg()` method.
3.  Pass as the aggregation function the string method to `.join`.

```{python shincol, eval = TRUE}

```

# Pandas Inventory

Now we know how to:

1.  Select columns

2.  Slice rows

3.  Do both simultaneously

4.  Filter rows using boolean masks

5.  Add columns

# Practice: Question 1

# Handy utility functions

**`sort_values()`**

```{python eval = FALSE}
uselections.sort_values("state")
```

**`value_counts()`**

```{python, eval = FALSE}
uselections["year"].value_counts()
```

**`unique()`**

```{python}

```

**`sample()`**

```{python}

```
