---
title: "Pandas II"
subtitle: "group by"
author: "Andrew Bray"
output: 
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE)
source("https://raw.githubusercontent.com/andrewpbray/stitched/master/demos/slides_css.R")
```

# Agenda

1.  Review
2.  Stuff
3.  Group by

---

# US Elections Data
--

-   From MIT Elections Lab
-   County-level Data
-   President elections 2000 - 2016

```{r r-setup}
# remotes::install_github("andrewpbray/boxofdata")
library(boxofdata)
library(tidyverse)
data(uselections)
dim(uselections)
names(uselections)
```

---

# US Elections Data, cont.
--

```{r, echo = FALSE}
select(uselections, -state_po, -FIPS, -office, -version) %>%
  head(8) %>%
  kable()
```


---

# Into Python
--

```{python}
import pandas as pd
uselections = r.uselections
uselections.shape
uselections.columns
uselections.dtypes
```


---

# Select columns

**Method 1:**  Pass a *string* into `[]`...

--

```{python select1, eval = FALSE}
uselections["county"]
```

--

```{python ref.label = "select1", echo = FALSE}
```

--

... get out a series.


---

# Select columns

**Method 2:** Pass a *list* into `[]`...

--

```{python select2, eval = FALSE}
uselections[["county"]]
```

--

```{python ref.label = "select2", echo = FALSE}
```

--

... get out a data frame.


---

# Slicing rows
--

**Method 1:** Pass a *slice* into `[]`...

--

```{python slice1, eval = FALSE}
uselections[0:5]
```

--

```{python ref.label = "slice1", echo = FALSE}
```

--

... get out a slice data frame. Sound familiar?

--

```{r}
slice(uselections, 1:2)
```


---

# Selecting and slicing
--

**Method 1:** (preferred) access labels with `.loc`.

```{python colandrow, eval = FALSE}
uselections.loc[0:5, ["county"]]
```

--

```{python ref.label = "colandrow", echo = FALSE}
```


---

# Selecting and slicing
--

**Method 2:** access integer indices with `.iloc`.

```{python checkcols, eval = FALSE}
uselections.columns
```

--

```{python ref.label = "checkcols", echo = FALSE}
```

--

```{python colandrow2, eval = FALSE}
uselections.iloc[0:5, 3]
```

--

```{python ref.label = "colandrow2", echo = FALSE}
```


---

# Filtering rows
--

You can apply a Boolean series as a mask.

```{python boolrows, eval = FALSE}
mask = uselections["year"] == 2016
uselections[mask]
```

--

```{python ref.label = "boolrows", echo = FALSE}
```


---

## Filtering rows and selecting columns

Boolean mask plus a list of columns.

```{python boolandcolsa, eval = FALSE}
mask = uselections["year"].isin([2012, 2016])
uselections[mask, ["county", "state"]]
```

Will this run?

--

> Need to use `.loc`

```{python boolandcolsb, eval = FALSE}
mask = uselections["year"].isin([2012, 2016])
uselections.loc[mask, ["county", "state"]] #<<
```

--

```{python ref.label = "boolandcolsb", echo = FALSE}
```


---

### Let's shine that up.

1.  Form data frame.
2.  Apply `.agg()` method.
3.  Pass as the aggregation function the string method to `.join`.

--

```{python shincol, eval = TRUE}

```

---

# Pandas Inventory

Now we know how to:

1. Select columns
--

2. Slice rows
--

3. Do both simultaneously
--

4. Filter rows using boolean masks
--

5. Add columns


---

# Practice: Question 1

---

# Practice: Question 1

Extract the first three rows where the candidate got more than 90% of the vote.

```{python q1}
uselections["prop"] = uselections["candidatevotes"]/uselections["totalvotes"]
```

---

# Handy utility functions

**`sort_values()`**

```{python eval = FALSE}
uselections.sort_values("state")
```

**`value_counts()`**

```{python, eval = FALSE}
uselections["year"].value_counts()
```

**`unique()`**

```{python}

```

**`sample()`**

```{python}

```


---

# Practice: Question 2


---

# Practice: Question 2

Which candidates were on the ballot in California in 2016?

```{python q2}
uselections["candidate"].unique()
mask = (uselections["year"] == 2016) & (uselections["state_po"] == "CA")
uselections[mask]["candidate"].unique()
```


---

# Practice: Question 3

Which were the top 5 counties in California in 2016 in the proportion of the vote won by Hillary Clinton?

```{python}
mask = (uselections["year"] == 2016) & (uselections["state_po"] == "CA") & (uselections["candidate"] == "Hillary Clinton")
df = uselections[mask]
df.sort_values("prop", ascending = False)["county"].head(5)
```


---

# Groupby
--

For separate operations on subsets of the data frame, use *grouped* operations.

```{python gb, eval = FALSE}
uselections.groupby("year")
```

--

```{python ref.label = "gb", echo = FALSE}

```

--

```{python gbtot, eval = FALSE}
uselections.groupby("year").agg(sum)
```

--

```{python ref.label = "gbtot", echo = FALSE}

```


---

# Practice: Question 4






