---
title: "Pandas II"
subtitle: "group by"
author: "Andrew Bray"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: atelier-forest-light
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE)
source("https://raw.githubusercontent.com/andrewpbray/stitched/master/demos/slides_css.R")
```


# Agenda

1. Review
2. Stuff
3. Group by


---
# US Elections Data
--

- From MIT Elections Lab
- County-level Data
- President elections 2000 - 2016

```{r r-setup}
# remotes::install_github("andrewpbray/boxofdata")
library(boxofdata)
library(tidyverse)
data(uselections)
dim(uselections)
names(uselections)
```

---
# US Elections Data, cont.

```{r, echo = FALSE}
select(uselections, -state_po, -FIPS, -office, -version) %>%
  head(8) %>%
  kable()
```

---
# Into Python
--

```{python}
import pandas as pd
uselections = r.uselections
uselections.shape
uselections.columns
uselections.dtypes
```


---
# Select columns
--

**Method 1:** Pass a *string* into `[]`...

--

```{python series, eval = FALSE}
uselections["county"]
```

--

```{python ref.label = "series", echo = FALSE}
```
--

... get out a series.


---
# Select columns

**Method 2:** Pass a *list* into `[]`...

--

```{python df, eval = FALSE}
uselections[["county"]]
```

--

```{python ref.label = "df", echo = FALSE}
```
--

... get out a dataframe.

---
# Slice rows
--

Pass a *slice* into `[]`...

```{python rows, eval = FALSE}
uselections[0:5]
```

--
```{python ref.label = "rows", echo = FALSE}
```
--

...get out a slice of rows. Does this terminology sound familiar?

--

```{r}
dplyr::slice(uselections, 1:2)
```


---
# Selecting and slicing
--

**Method 1:** (preferred) access labels with `.loc`.

```{python colandrow, eval = FALSE}
uselections.loc[0:5, ["county"]]
```

--
```{python ref.label = "colandrow", echo = FALSE}
```


---
# Selecting and slicing
--

**Method 2:** access integer indices with `.iloc`.

```{python checkcols, eval = FALSE}
uselections.columns
```

--
```{python ref.label = "checkcols", echo = FALSE}
```

--

```{python colandrow2, eval = FALSE}
uselections.iloc[0:5, 3]
```

--
```{python ref.label = "colandrow2", echo = FALSE}
```

---
# Filtering rows
--

You can apply a Boolean series as a mask.

```{python boolrows, eval = FALSE}
mask = uselections["year"] == 2016
uselections[mask]
```
--

```{python ref.label = "boolrows", echo = FALSE}
```


---
# Filtering rows

FYI: use `.isin` to mask based on set membership.

```{python isin, eval = FALSE}
mask = uselections["year"].isin([2012, 2016])
uselections[mask]
```
--

```{python ref.label = "isin", echo = FALSE}
```


---
# Filtering rows and selecting columns
--

Boolean mask plus a list of columns.

```{python boolandcolsa, eval = FALSE}
mask = uselections["year"].isin([2012, 2016])
uselections[mask, ["county", "state"]]
```

Will this run?

--

> Need to use `.loc`

```{python boolandcolsb, eval = FALSE}
mask = uselections["year"].isin([2012, 2016])
uselections.loc[mask, ["county", "state"]] #<<
```
--

```{python ref.label = "boolandcolsb", echo = FALSE}
```

---
# Adding a column
--







