---
title: "Pointers, Equality, and Identity"
output: github_document
---


## Pointers in R

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(reticulate)
```

Let's start by creating a vector of the integers 1 to 10 in R.

```{r}
x <- 1:10
x
```

We can ask for the location of this object in memory using `obj_addr()`.

```{r}
library(lobstr)
obj_addr(x)
```

Let's point another variable `y` at `x` and see where it is stored.

```{r}
y <- x
obj_addr(y)
```

We can see that's they're actually pointing to the same object. This is a sensible thing to do to save memory - why copy over something that's the same?

Let's see what happens if we if we change `x` in place:

```{r}
x[1] <- 99
obj_addrs(list(x, y))
```

We see that while `y` is still pointing to the original object, R has copied over the contents of `x` into a new location, then made the change. This behavior is helpful if you want to be sure to not inadvertently overwrite objects, but it does make me wonder why we call that syntax "change in place"! Perhaps "change in object" would be better.


## Pointers in Python

Let's go through the same process with Python.

```{python}
x = list(range(1, 11))
x
y = x
[id(x), id(y)]
```

So far they're behaving the same way: both `x` and `y` are pointing to the same object in memory. Let's see what happens when we change `x`.

```{python}
x[0] = 99
[id(x), id(y)]
```

*That's* different. Python truly does change `x` in place; it maintains the object's location in memory so that anything else also pointed to that memory slot is also changed.


## Equality and Identity

Two objects are *equal* if they share the same value. Two objects are *identical* if they're actually the same object in memory. In Python, `x` and `y` are both identical and equal.

```{python}
x is y # identical
x == y # equal
```

We can create a third variable `z`, that's equal,

```{python}
z = [99, 2, 3, 4, 5, 6, 7, 8, 9, 10]
z == x
```

But because it's not referring to the same object in memory,  it's not identical.

```{python}
[id(z), id(x)]
z is x
```

In R, `x` and `y` were both equal and identical initially, but when we changed `x`, we broke both the equality and the identity.

```{r}
x == y # not equal
obj_addr(x) == obj_addr(y) # not identical
```

Note that there is a function called `identical()` in R that isn't using the same definition of *identical* that we're using here. Instead of outputting a vector of `TRUE` and `FALSE` for each pairwise logical comparison that we're making between `x` and `y`, it only returns `TRUE` if all values are `TRUE` and `FALSE` otherwise

```{r}
identical(x, y)
```



